---
title: "mixsiar_script_lakes"
author: "Jonathan Raberg"
date: "4/27/2018"
output:
  html_document: 
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float: true
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

*Iceland lakes that could work:*
* ARN (between algae and aquatic plants)
* HVT (though it has some outside isospace)
* TORF (between algae and aquatic plants)
* TRK (between a, ap, and t!)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Brian Stock
March 8, 2016
Script file to run wolves example without GUI

# Load MixSIAR package
```{r}
library(tidyverse)
library(MixSIAR)
library(latex2exp)
library(gridExtra)
library(grid)
library(lattice)
```

# Initialize

```{r}
# Update the lake name and run length, create directory
lake_name <- "HAK" # for plots
lake <- "hak_37" # for use in code

run_length <- "long"

my_dir <- file.path("suess_corrected", paste(lake, run_length, sep = "_")) # add appendix here if needed
if(!dir.exists(my_dir)) {
  dir.create(my_dir)
} else {
  "Error: Directory already exists"
}


options(max.print = 1000) # Their default
```
## Reload previous data

```{r}
# load(file.path(my_dir, "jags1.saved"))
```



# Load mixture data
i.e. your:
  -Consumer isotope values (trophic ecology / diet)
  -Mixed sediment/water tracer values (sediment/hydrology fingerprinting)

 'filename' - name of the CSV file with mix/consumer data
 'iso_names' - column headings of the tracers/isotopes you'd like to use
 'random_effects' - column headings of any random effects
 'cont_effects' - column headings of any continuous effects

 'iso_names', 'random_effects', and 'cont_effects' can be a subset of your columns
   i.e. have 3 isotopes in file but only want MixSIAR to use 2,
   or have data by Region and Pack but only want MixSIAR to use Region

To run on your data, replace the system.file call with the path to your file

## Lake data

```{r}
# To run on your data, replace the system.file call with the path to your file
mix.filename <- file.path(paste("extdata/", lake, "_sed.csv", sep = ""))

# Load mixture data
mix <- load_mix_data(filename=mix.filename, 
                     iso_names=c("d13C","d15N"), 
                     factors=c("Age"), 
                     fac_random=c(TRUE), 
                     fac_nested=c(FALSE), 
                     cont_effects=NULL)
```


# Load source data
i.e. your:
  -Source isotope values (trophic ecology / diet)
  -Sediment/water source tracer values (sediment/hydrology fingerprinting)

 'filename': name of the CSV file with source data
 'source_factors': column headings of random/fixed effects you have source data by
 'conc_dep': TRUE or FALSE, do you have concentration dependence data in the file?
 'data_type': "means" or "raw", is your source data as means+SD, or do you have raw data

## Lake sources

```{r}
# To run on your data, replace the system.file call with the path to your file
source.filename <- file.path("extdata", "iceland_sources_a_split_suess.csv")
# file.exists("extdata/qpt_sources.csv")

# Load source data
source <- load_source_data(filename=source.filename, source_factors=NULL, 
                            conc_dep=FALSE, data_type="means", mix)
```


#Load discrimination data
i.e. your:
   -Trophic Enrichment Factor (TEF) / fractionation values (trophic ecology/diet)
   -xxxxxxxx (sediment/hydrology fingerprinting)

'filename' - name of the CSV file with discrimination data

## Lakes discriminations (= 0)

```{r}
# To run on your data, replace the system.file call with the path to your file
discr.filename <- file.path("extdata", "discrimination.csv")
# file.exists("extdata/wolves_discrimination.csv")

# Load discrimination data
discr <- load_discr_data(filename=discr.filename, mix)
```


# Make isospace plot
Are the data loaded correctly?
Is your mixture data in the source polygon?
Are one or more of your sources confounded/hidden?

 'filename' - name you'd like MixSIAR to save the isospace plot as 
              (extension will be added automatically)
 'plot_save_pdf' - TRUE or FALSE, should MixSIAR save the plot as a .pdf?
 'plot_save_png' - TRUE or FALSE, should MixSIAR save the plot as a .png?
 
```{r}
# plot_data(filename=file.path(my_dir, "isospace_plot"), 
#           plot_save_pdf=TRUE,
#           plot_save_png=FALSE,
#           mix,source,discr) 
```

## My isospace plot


```{r, warning=FALSE}
lake_data <- read_csv(mix.filename)
source_data <- read_csv(source.filename)

isospace <- 
  ggplot() + 
  geom_point(data = lake_data, aes(x = d13C, y = d15N, color = Age)) + 
  #geom_path(data = lake_data, aes(x = d13C, y = d15N, color = Age), size = 0.2) +
  scale_color_gradient2(low = "dark blue", mid = "dark green", high = "gold", midpoint = 3500) +
  geom_point(data = source_data, aes(x = Meand13C, y = Meand15N), color = "black") + 
  geom_errorbar(data = source_data, aes(
    x = Meand13C, y = Meand15N, ymin = Meand15N - SDd15N, ymax = Meand15N + SDd15N)) +
  geom_errorbarh(data = source_data, aes(
    x = Meand13C, y = Meand15N, xmin = Meand13C - SDd13C, xmax = Meand13C + SDd13C)) +
  geom_text(data = source_data, aes(x = Meand13C, y = Meand15N, label = Source, hjust = -0.5, vjust = 2)) +
  labs(
    title = paste("Lake", lake_name, "Isospace"), 
    color = "Age (yrs)", 
    x = latex2exp::TeX("$\\delta^{13}C$"), 
    y = latex2exp::TeX("$\\delta^{15}N$"))
isospace
ggsave(file.path(my_dir,"isospace.pdf"), plot = isospace, width = 8, height = 5)

```

If 2 isotopes/tracers, calculate normalized surface area of the convex hull polygon(s)
   *Note 1: discrimination SD is added to the source SD (see calc_area.r for details)
   *Note 2: If source data are by factor (as in wolf ex), computes area for each polygon
             (one for each of 3 regions in wolf ex)

```{r}
if(mix$n.iso==2) calc_area(source=source,mix=mix,discr=discr)
```

Define your prior, and then plot using "plot_prior"
   RED = your prior
   DARK GREY = "uninformative"/generalist (alpha = 1)
   LIGHT GREY = "uninformative" Jeffrey's prior (alpha = 1/n.sources)

 default "UNINFORMATIVE" / GENERALIST prior (alpha = 1)
 
```{r}
plot_prior(alpha.prior=1,source, filename = file.path(my_dir, "prior_plot"))
dev.off()
```

# Write JAGS model file (define model structure)
Model will be saved as 'model_filename' ("MixSIAR_model.txt" is default,
    but may want to change if in a loop)

There are 3 error term options available:
   1. Residual * Process (resid_err = TRUE, process_err = TRUE)
   2. Residual only (resid_err = TRUE, process_err = FALSE)
   3. Process only (resid_err = FALSE, process_err = TRUE)

 'model_filename': don't need to change, unless you create many different models
 'resid_err': include residual error in the model?
 'process_err': include process error in the model?

  *Note: If you have only 1 mix datapoint, you have no information about the 
         mixture/consumer variability. In this case, we ues the original MixSIR
         error model (which does not fit a residual error term).
         This is the same behavior as 'siarsolo' in SIAR.

```{r}
model_filename <- file.path("models", "MixSIAR_model.txt")
resid_err <- TRUE
process_err <- TRUE
write_JAGS_model(model_filename, resid_err, process_err, mix, source)
```

# Run model
JAGS output will be saved as 'jags.1'

 MCMC run options:
 run <- "test"       # chainLength=1000, burn=500, thin=1, chains=3, calcDIC=TRUE
 run <- "very short" # chainLength=10000, burn=5000, thin=5, chains=3, calcDIC=TRUE
 run <- "short"      # chainLength=50000, burn=25000, thin=25, chains=3, calcDIC=TRUE
 run <- "normal"     # chainLength=100000, burn=50000, thin=50, chains=3, calcDIC=TRUE
 run <- "long"       # chainLength=300000, burn=200000, thin=100, chains=3, calcDIC=TRUE
 run <- "very long"  # chainLength=1000000, burn=500000, thin=500, chains=3, calcDIC=TRUE
 run <- "extreme"    # chainLength=3000000, burn=1500000, thin=500, chains=3, calcDIC=TRUE

 Can also set custom MCMC parameters
 run <- list(chainLength=200000, burn=150000, thin=50, chains=3, calcDIC=TRUE)

## Real run

```{r}
# After a test run works, increase the MCMC run to a value that may converge
jags.1 <- run_model(run=run_length, mix, source, discr, model_filename,
                    alpha.prior = 1, resid_err, process_err)
```


# Process JAGS output

## Save jags.1

```{r}
save(jags.1, file = file.path(my_dir, "jags1.saved"))
```

## Parse jags.1

For documentation, see p. 22 of https://github.com/brianstock/MixSIAR/blob/master/Manual/mixsiar_manual_3.1.pdf


```{r}
# Pull ages. This is really stupid, but used bind_rows to stack the ages, once for each source type
ages <- read_csv(mix.filename)["Age"]
ages <- bind_rows(ages, ages, ages, ages)

# Pull means
means <- as_tibble(jags.1[["BUGSoutput"]][["mean"]][["p.fac1"]]) %>% 
  rename(
    "Algae" = V1,
    "Aquatic_plants" = V2,
    "Soil" = V3,
    "Terrestrial" = V4
  ) %>%
  gather(Type, Mean, "Algae", "Aquatic_plants", "Soil", "Terrestrial")

# Pull standard deviations
sd <- as_tibble(jags.1[["BUGSoutput"]][["sd"]][["p.fac1"]]) %>% 
  rename(
    "Algae" = V1,
    "Aquatic_plants" = V2,
    "Soil" = V3,
    "Terrestrial" = V4
  ) %>% 
  gather(Type, SD, "Algae", "Aquatic_plants", "Soil", "Terrestrial")

# Pull medians
medians <- as_tibble(jags.1[["BUGSoutput"]][["median"]][["p.fac1"]]) %>% 
  rename(
    "Algae" = V1,
    "Aquatic_plants" = V2,
    "Soil" = V3,
    "Terrestrial" = V4
  ) %>% 
  gather(Type, Median, "Algae", "Aquatic_plants", "Soil", "Terrestrial")

sum_stats <- bind_cols(ages, means["Type"], means["Mean"], sd["SD"], medians["Median"])
```

## Plots from jags.1

```{r}
p <- sum_stats %>%
  ggplot() +
  aes(Age, Mean) +
  scale_x_reverse() +
  labs(title = paste("Lake", lake_name), subtitle = paste("Run length =", run_length))

means_area_plot <- p +
  aes(fill = Type) +
  geom_area()

medians_area_plot <- p +
  aes(x = Age, y = Median, fill = Type) +
  geom_area()

means_point_plot <- p +
  aes(color = Type) +
  geom_point() +
  geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=200) +
  labs(y = "Mean +/- SD")


means_area_plot
medians_area_plot
means_point_plot


# # If you want SD included in the area plot, start here...
# sum_plot <- sum_stats_mod %>%
#   mutate(
#     Mean = Mean - SD
#   ) %>% 
#   gather(var, value, Mean, SD) %>% 
#   mutate(
#     Type = paste(Type, var, sep = "_")
#     ) %>% 
#   ggplot() +
#   aes(Age, value, fill = Type) +
#   #geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.1) + # This puts the bars in the wrong spots
#   geom_area()
# sum_plot
```


```{r}
# Save plots
ggsave(file.path(my_dir,"means_area_plot.pdf"), plot = means_area_plot, width = 8, height = 5)
ggsave(file.path(my_dir,"medians_area_plot.pdf"), plot = medians_area_plot, width = 8, height = 5)
ggsave(file.path(my_dir,"means_point_plot.pdf"), plot = means_point_plot, width = 8, height = 5)
ggsave(file.path(my_dir,"means_meds_plots_together.pdf"), plot = grid.arrange(means_area_plot, medians_area_plot, means_point_plot))
```


## JAGS output

Choose output options (see ?output_options for details)

### Diagnostics

```{r}
# Nice to suppress all plots the first time to see 
diagnostics <- TRUE
plots <- FALSE


# "sup" options must be false for plots to print, for some reason
output_options <- list(summary_save = diagnostics,                 
                       summary_name = file.path(my_dir, "summary_statistics"), 
                       sup_post = !plots,                    
                       plot_post_save_pdf = plots,           
                       plot_post_name = file.path(my_dir, "posterior_density"),
                       sup_pairs = !plots,             
                       plot_pairs_save_pdf = plots,    
                       plot_pairs_name = file.path(my_dir, "pairs_plot"),
                       sup_xy = !plots,           
                       plot_xy_save_pdf = plots,
                       plot_xy_name = file.path(my_dir, "xy_plot"),
                       gelman = diagnostics,
                       heidel = FALSE,  
                       geweke = diagnostics,   
                       diag_save = diagnostics,
                       diag_name = file.path(my_dir, "diagnostics"),
                       indiv_effect = FALSE,       
                       plot_post_save_png = FALSE, 
                       plot_pairs_save_png = FALSE,
                       plot_xy_save_png = FALSE)
```

```{r}
# Check to see if plots will work
dev.list() # This shows current devices. If above 60, R won't work.
while (!is.null(dev.list()))  dev.off()
```


```{r}
# Create diagnostics, summary statistics, and posterior plots
options(max.print = 99999)
output_JAGS(jags.1, mix, source, output_options)
```

### Density plots

```{r}
# Nice to suppress all plots the first time to see 
diagnostics <- FALSE
plots <- TRUE


# "sup" options must be false for plots to print, for some reason
output_options <- list(summary_save = diagnostics,                 
                       summary_name = file.path(my_dir, "summary_statistics"), 
                       sup_post = !plots,                    
                       plot_post_save_pdf = plots,           
                       plot_post_name = file.path(my_dir, "posterior_density"),
                       sup_pairs = !plots,             
                       plot_pairs_save_pdf = plots,    
                       plot_pairs_name = file.path(my_dir, "pairs_plot"),
                       sup_xy = !plots,           
                       plot_xy_save_pdf = plots,
                       plot_xy_name = file.path(my_dir, "xy_plot"),
                       gelman = diagnostics,
                       heidel = FALSE,  
                       geweke = diagnostics,   
                       diag_save = diagnostics,
                       diag_name = file.path(my_dir, "diagnostics"),
                       indiv_effect = FALSE,       
                       plot_post_save_png = FALSE, 
                       plot_pairs_save_png = FALSE,
                       plot_xy_save_png = FALSE)
```


```{r}
# Create diagnostics, summary statistics, and posterior plots
options(max.print = 99999)
output_JAGS(jags.1, mix, source, output_options)
```